---
title: "CleaningCounties"
author: "Kelley Erickson"
date: "April 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

As a working example, let's look at records of Tracheophytes in Florida. These data came from downloading  (on GBIF) all of the occurrences of Tracheophytes that: 
1. Had no coordinates
2. stateProvince == "Florida"

The shape polygon 'florida' that I loaded in came from the GADM shape file for the United States clipped to Florida. 
```{r, load data}

load("C:/Users/kerickson/Documents/GitHub/!Data/Tracheophytes No Coordinates/tracheophytes_nocoords.RData")

load("C:/Users/kerickson/Documents/GitHub/!Data/florida.RData")

head(tracheophytes_nocoords)
```

Now let's get the list of county names from GADM as well as all the different ways that counties are written in our dataset from GBIF: 
```{r, getting list of GADM county names}

#What are the GADM names of the counties in Florida? 
GADM_counties <- unique(florida@data$NAME_2)
#Refactor so levels only correspond to counties in FL (optional)
GADM_counties <- factor(GADM_counties) 
GADM_counties 
length(GADM_counties) 

#There are 67 counties in Florida 

length(GADM_counties) 
```

And now the GBIF counties: 
```{r, getting list of GBIF county names} 

GBIF_counties <- unique(tracheophytes_nocoords$county)
#Refactor so levels only correspond to counties in FL (optional)
GBIF_counties <- factor(GBIF_counties)
GBIF_counties 

length(GBIF_counties)
```

Notice how in the GBIF_counties: 

 * Some of the counties have state and country: Franklin County, Florida, United States
 * Some of them are written in all-caps: OSCEOLA vs. Osceola
 * Some of them have parentheses around them: (Lake) instead of Lake 
 * Some of them have the state in parentheses: Leon Co. (FL)
 * A single county could be written multiple ways: Fulton Co., Fulton, Fulton County
 * There are misspellings: Semiole (should be Seminole) 
 * There are also quite a few that don't seem to match up with any of the GADM  counties
 
 If we were to only use records that had county names that matched GADM-county names we would be throwing out a lot of records: 
 
```{r, see how many records actually match GADM counties right now}


tracheophytes_clean <- subset(tracheophytes_nocoords, tracheophytes_nocoords$county %in% GADM_counties)
nrow(tracheophytes_clean)

#Now there are 128,268 records with GADM standard names 
```
#1. Remove extraneous text 
First let's focus on removing all of the extraneous punctuation and words. I have put several lines of code into a function so that it can be easily applied to different data frames.   

```{r, remove extraneous text} 


removeExtraText <- function(df_column) {
  df_column <- gsub(df_column, pattern='(', replacement='', fixed=T)
  df_column <- gsub(df_column, pattern=')', replacement='', fixed=T)
  df_column <- gsub(df_column, pattern=' County', replacement='', 
                            fixed=T)
  df_column <- gsub(df_column, pattern = ', Florida, United States', replacement = "", fixed=T)
  df_column <- gsub(df_column, pattern =' COUNTY', replacement='', fixed=T)
  df_column <- gsub(df_column, pattern='County of ', replacement='')
  df_column <- gsub(df_column, pattern=' Cty.', replacement='')
  df_column <- gsub(df_column, pattern=' Cty', replacement='')
  df_column <- gsub(df_column, pattern=' Co.', replacement='')
  df_column <- gsub(df_column, pattern='Ste ', replacement='Saint ')
  df_column <- gsub(df_column, pattern='St. ', replacement='Saint ')
  df_column <- gsub(df_column, pattern='St ', replacement='Saint ')
  df_column <- gsub(df_column, pattern ='ST.', replacement = 'Saint')
  df_column <- gsub(df_column, pattern = 'st.', replacement = 'Saint')
  df_column <- gsub(df_column, pattern = " ' ", replacement = "")
  df_column <- gsub(df_column, pattern='[', replacement='', fixed=T)
  df_column <- gsub(df_column, pattern=']', replacement='', fixed=T)
  df_column <- gsub(df_column, pattern=' FL', replacement='')
return(df_column)
} 


tracheophytes_nocoords$county <- removeExtraText(tracheophytes_nocoords$county)

#Let's check what we've fixed so far: 
unique(tracheophytes_nocoords$county)

tracheophytes_clean <- subset(tracheophytes_nocoords, tracheophytes_nocoords$county %in% GADM_counties)
nrow(tracheophytes_clean)

#Now there are 240, 175 records with GADM standard county names. An improvement!
```  
Starting to look better!
#2. Fix capitalization 
The next thing to standardize is how GBIF counties are capitalized: 
```{r, fix capitalization}
#The capwords function comes from the chartr {base} helpfile

#First convert everything to lowercase 

tracheophytes_nocoords$county <- tolower(tracheophytes_nocoords$county)

#Now recapitalize using sentence case using a function that was written out in the chartr {base} helpfile 

capwords <- function(s, strict = FALSE) {
  cap <- function(s) paste(toupper(substring(s, 1, 1)),
                           {s <- substring(s, 2); if(strict) tolower(s) else s},
                           sep = "", collapse = " " )
  sapply(strsplit(s, split = " "), cap, USE.NAMES = !is.null(names(s)))
}

tracheophytes_nocoords$county <- capwords(tracheophytes_nocoords$county)
#Check our work: All counties should now be in regular sentence case 
unique(tracheophytes_nocoords$county) 

tracheophytes_clean <- subset(tracheophytes_nocoords, tracheophytes_nocoords$county %in% GADM_counties)
nrow(tracheophytes_clean)
#Now there are 225, 821 records that have GADM-standard counties. Not quite sure why it went down again...  
```

 At this point in the cleanup, the main thing we notice about the list of GBIF county names is that there are a lot of misspellings. This is the part that involves a lot of looking at differences.
 
 I have written two functions `compare` and `printIssues` that help automate some of the process. First the `compare` function figures out all of the counties whose names are "close-enough" to other counties names 
 
 The `printIssues` function prints out groupings that the `compare` function noticed as being potentially similar to eachother. Groupings are separated by '...'. 
 
 You will notice right away that not everything the computer identifies as being similar is actually similar. Hence the need for a human to read through them and spot similarities. 

```{r, a function to help spot similar looking spellings} 
#The function compare 
compare <- function(this, that) {
    issues <- NULL

    for (i in 1:length(this)){
      test<-agrep(this[i], that)
      if (length(test)>1) {
        issues <- c(issues, as.character(this[i]))
      }
    }

return(issues)
}

#The function printIssues 
printIssues <- function (issues, compareList) {
  for (i in 1:length(issues)) {
  print(agrep(issues[i], compareList, value=T))
  print("...")
  }
}

counties_GBIF <- unique(tracheophytes_nocoords$county)
#These functions will not work if there is a blank county name: ""
counties_GBIF <- subset(counties_GBIF, counties_GBIF !="")

issues<- compare(counties_GBIF, counties_GBIF)
printIssues(issues, counties_GBIF)
```

#The final step (really it's an iterative process):
Now let's fix some of those spellings by making use of a crosswalk table. I have gone through the list above and put misspellings that I noticed into the csv file. 

As you go through you will probably introduce some new errors: For example if you are trying to convert 'Saint John' to the GADM-standard Saint Johns, records that already had Saint Johns correctly spelled will suddenly become Saint Johnss. So you will then need to replace 'Saint Johnss' with Saint Johns. So it's a very iterative process.

The .csv file counties_lookup contains the crosswalk table I constructed. The function `renameCounties` makes use of the crosswalk table to correct the spelling of GBIF counties to resemble GADM spellings. I have so far only been working with counties in Florida. This function and the ones above will have to be modified to work for different states. 

```{r, making use of the crosswalk table}

renameCounties <- function(df) {

  temp <- read.csv(file="C:/Users/kerickson/Documents/GitHub/sundries/R/county_lookup.csv"  )
  for(i in 1:nrow(temp)) {
    df$recordedBy <- gsub(df$recordedBy, pattern = paste0("\\b", temp$Incorrect[i], "\\b"), 
                          replacement = temp$GADM[i], fixed=T)
  }

  return(df)
}

tracheophytes_nocoords <-renameCounties(tracheophytes_nocoords)
tracheophytes_clean <- subset(tracheophytes_nocoords, tracheophytes_nocoords$county %in% GADM_counties)
nrow(tracheophytes_clean)
#Now there are 225, 822 records that have GADM-standard counties. Not quite sure why it went down again...
tracheophytes_messy <- subset(tracheophytes_nocoords, !(tracheophytes_nocoords$county %in% GADM_counties))

unique(tracheophytes_messy$county)


```
#Next step: figure out why the renameCounty function isn't working 100%

As you can see from the unique list of "messy" records (those whose county still doesn't officially match a GADM county), there are still some counties whose spellings _should_ have been fixed by the renameCounty function. And yet... 

I modified this workflow from a similar process for renaming collectors, so haven't spent as much time correcting and investigating the county side of things, but I will continue to look into it and update this site as I figure things out. 
